Structuring College Access: The Market Segment Model and College Board Geomarkets

### 2025-04-25 text/back_end_KS.qmd
rq2 LA results

@@

```{r fig-ls-rq2-midSAT-combo, echo = FALSE, fig.align = 'center', fig.cap = "Los Angeles Geomarket Contribution to Purchased Student profiles by First-Generation Status, Middle-Range SAT orders", out.width = "100%"}
knitr::include_graphics('./../assets/images/rq2_la_midSAT_combo.png')
#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
#![The enrollment funnel](assets/images/enrollment_funnel.png)
```

>>>>

We assess how the socioeconomic and racial composition of included versus excluded prospects varies when student list purchases filter on particular Geomarkets in the Los Angeles metropolitan areas by analyzing six orders placed by a research university. These six orders filtered for prospects across the entire state of California, in the 2019-2020 high school graduating class, and by PSAT scores. Two orders indicated PSAT score thresholds ranging from 1070-1180, another two orders filtered for a 1190-1260 PSAT range, and the remaining two orders filtered for scores ranging from 1270-1520.

The racial/ethnic composition of prospects in the Los Angeles metropolitan area whose student profiles were purchased across these orders is presented in @fig-la-rq2-race by Geomarket. The more than 15,000 prospects in the resulting student lists at the low PSAT score thresholds (1070-1180) reflect a very similar racial/ethnic composition within Geomarkets than the overall demographic makeup of the metropolitan area in 2020 (see @fig-la-rq1-race). However, the racial/ethnic composition of prospects become less proportional to the overall metropolitan by Geomarket as PSAT scores increase. For instance, prospect pools at middle range PSAT scores (1190-1260) within Geomarkets become disproportionately more Asian, while the representation of Hispanic and Black prospects decline substantially. This pattern becomes most pronounced at the highest PSAT score prospect pools. This disproportionate over representation for Asian prospects in comparison to the population of residents with Geomarkets ranges from 3 percentage points in the West Los Angeles and West Beach Geomarket (16% prospects versus 13% Asian residents) to 52 percentage points in the East Los Angeles Geomarket (79% prospects versus 27% residents). @fig-la-rq2-race illustrates the share of Black prospects within these high PSAT orders declines to less than 3% within all Geomarkets in comparison to Black residents making up a range of 3% (Covina and West Covina) to 18% (South and South Central Los Angeles) of all residents within Geomarkets. Similar underrepresentation patterns are evident for Hispanic prospects across all Geomarkets execept for a nearly equal representation in South and South Central Los Angeles (73% prospects versus 71% Hispanic residents).

@fig-la-rq2-midSAT-race presents the race/ethnicity of all prospects targeted in the middle PSAT score range across Geomarket contributions for the Los Angeles metropolitan area. For the more than 2,500 White prospects in this pool, we see the San Fernando Valley - West (28% versus 20% for all prospects) and the West Los Angeles and West Beach (12% versus 7% for all prospects)Geomarkets contributing disproportionately larger shares. Larger disparities across Geomarket contributions were evident for Asian, Black, and Hispanic students. The Glendale and Pasadena (17% versus 14% for all prospects), East Los Angeles (11% versus 5% for all prospects) and Covina and West Covina (17% versus 11% for all prospects) Geomarkets contributed a disproportionate larger share of Asian students, whereas nearly all other Geomarkets contributed disproportionately smaller or nearly equal shares.

Representational patterns across Geomarkets are reversed for Black and Hispanic prospects in @fig-la-rq2-midSAT-race. South and South Central Los Angeles and Riverside, San Bernadino, and Ontario Geomarkets contributed larger proportions of Black and Hispanic prospects relative to their contributions to the overall prospect pool in the Los Angeles metropolitan area. On the other hand, the San Fernando Valley- West and the Glendale and Pasadena Geomarket contributed smaller proportions of Black and Hispanic Prospects. However, the West Los Angeles and West Beach (10% versus 7% for all prospects), Hollywood Wilshire (7% versus 5% for all prospects), and the Long Beach (8% versus 5% for all prospects) Geomarkets contributed a larger share of Black prospects, whereas Covina and West Covina contributed a larger share of Hispanic students. All other Geomarkets contributed nearly equal shares of Black and Hispanic prospects.

Contributions of each Geomarket to the pool of prospects from the Los Angeles metropolitan area based on first-generation college student status are presented in @fig-la-rq2-midSAT-ses. Only 889 of the more than 8,000 total prospects (11%) in this middle PSAT score range (1190-1260) pool were first-generation college students whose parents did not attend college. Only three of the 11 Geomarkets contributed substantially larger shares of these first-generation prospects: East Los Angeles (35%), South and South Central Los Angeles (47%), and Riverside, San Bernardino, and Ontario (15%). However, disproportionate contributions become more spread across Geomarkets for the nearly 1,436 of the more than 8,000 total prospects (18%) who were first-generation college students whose parents attended but did not complete their degree. Overall, six of the 11 Geomarkets contributed larger shares of these prospects: Hollywood & Wilshire (25%), East Los Angeles (26%), South and South Central Los Angeles (28%), Long Beach (25%), Covina and West Covina (22%), and Riverside, San Bernardino, and Ontario (22%). This leads to the remaining Geomarkets - San Fernando Valley- West (80%), San Fernando Valley- East (78%), Glendale and Pasadena (77%), West Los Angeles & West Beach (87%), South Bay (81%) - contributing disproportionately larger shares of not first-generation college students relative to the overall prospect pool (71%).

Lastly, @fig-la-rq2-midSAT-combo presents Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity in the Los Angeles metropolitan area. Among all White prospects (n=2,511) whose contact information was purchased in this middle PSAT score range, approximately 3% were first-generation college students whose parents did not attend college, 11% were first-generation college students whose parents attended some college but did not complete their degree, and 86% were not first-generation college students. @fig-dla-rq2-midSAT-combo shows all Geomarkets contribute a nearly proportionate share of prospects by first-generation college students (across both no college and some college parents) relative to the overall pool. The South and South Central Los Angeles Geomarket stands out as one exception contributing a relatively larger share of White first-generation college students whose parents did not attend college (17%) and whose parent attended some college but did not complete their degree (33%). Other Geomarket that contributed signifcantly larger shares of White first-generation prospects whose parents did not complete their degree include Long Beach (17%), and Riverside, San Bernardino, and Ontario (19%).

Geomarkets contributions to first-generation college students was mixed across other racial/ethnic categories in @fig-la-rq2-midSAT-combo. For Asian prospects (11% first-generation no college, 21% first-generation some college, and 68% not first-generation), the Hollywood & Wilshire (12%) and East Los Angeles (36%) Geomarkets contributed larger proportions of first-generation college students whose parents did not attend college. The San Fernando Valley- East (28%), Hollywood & Wilshire (35%), East Los Angeles (25%), South and South Central Los Angeles (44%), Long Beach (27%), and Covina and West Covina (25%) Geomarkets also contributed larger proportions of first-generation college students whose parents attended college but did not complete their their degree.

For Black prospects (5% first-generation no college, 20% first-generation some college, and 74% not first-generation), Geomarkets that disproportionately larger proportions of first-generation college students whose parents did not attend college include South Bay (11%), South and South Central Los Angeles (21%), Long Beach (7%), and Covina and West Covina (9%). The West Los Angeles and West Beach (26%), Hollywood & Wilshire (36%), East Los Angeles (33%), Long Beach (40%), and Covina and West Covina (27%) Geomarkets also contributed larger proportions of Black first-generation college students whose parents attended college but did not complete their their degree.

Finally, Hispanic prospects in this middle range PSAT order were on average 25% first-generation college students whose parents did not attend college, 25% first-generation college students whose parents attended some college but did not complete their degrees, and 50% were not first-generation college students. The Hollywood & Wilshire (30%), East Los Angeles (39%), South and South Central Los Angeles (54%), and Riverside, San Bernardino, and Ontario (31%) contributed disproportionate shares of Hispanic first-generation college students whose parents did not attend college. The East Los Angeles (33%), South and South Central Los Angeles (28%), Long Beach (30%), and Riverside, San Bernardino, and Ontario (27%) also contributed larger Hispanic first-generation college students whose parents attended college but did not complete their their degree.

@@

```{r fig-la-rq2-midSAT-combo, echo = FALSE, fig.align = 'center', fig.cap = "Los Angeles Geomarket Contribution to Purchased Student profiles by First-Generation Status, Middle-Range SAT orders", out.width = "100%"}
knitr::include_graphics('./../assets/images/rq2_la_midSAT_combo.png')
#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
#![The enrollment funnel](assets/images/enrollment_funnel.png)
```
